Allee Effect and Cultural Evolution

### 2015-06-08 manuscript/draft.tex
Updates

When the magnitude of social learning is sufficiently high, the system is subject to effects similar to an increase in the per capita growth rate $r$, with the consequent decrease in the stability of the fixed points. Figure \ref{fig:bifurcationWithTransmission} shows the bifurcation diagram for three different initial settings of $a$ and $b$. In the first two cases, increase in $z$ leads to a shift of the equilibria, which remains stable even when $z$ is increased. The third case shows, around $z=0.5$ a transition from a point attractor to a system characterised by repeated shifts between the two transmission strategies.

%This should be changed with the basin plot and the time-series; omit discussion on bifurcation

\begin{figure*}
  \centering
      \includegraphics[width=1\textwidth]{./figures/figure4.jpg}
  \caption{Bifurcation diagram for $a$ (red) and $b$ (blue) as a function of social learning ($z$): left: $a=0.11$, $b=0.4$; Middle: $a=0.5$, $b=0.21$, $a=0.4$, $b=0.4$. Model parameters:  $A_a=0.1$; $K_a=0.6$; $A_b=0.2$; $K_b=0.7$; $r_a=r_b=0.05$}
    \label{fig:bifurcationWithTransmission}
\end{figure*}

High reliance on social learning thus generates a new basin of attraction ... %Not sure how to link this to the unstability figures. I think we already made the point with the analysis above, and I'm no longer sure what this new plot is ultimately adding (I know, I asked you to make it, sorry!)

Small differences in initial conditions can have major consequences in the boundaries between the unstable region and the rest of equilibria. Figure \ref{fig:unstable} shows an example with high value of social learning $z=0.4$, where two scenarios ($j$ and $k$) generate different outcomes from very close initial conditions.

%\begin{figure}
%  \centering
%      \includegraphics[width=0.5\textwidth]{./figures/figureUnstable1.pdf}
%  \caption{}
%    \label{fig:bifurcationWithTransmission}
%\end{figure}

>>>>

When the magnitude of social learning is sufficiently high, the system is subject to effects similar to an increase in the per capita growth rate $r$, with the consequent decrease in the stability of the fixed points. Figure XXXX1 shows how an increase in $z$ generate time-series with shifts from one variant to the other. This regime is prone to occur when both variants are initialised with high densities (XXXX2/basin)/. In these conditions the total density is too high to ensure a stable coexistence, and the regulatory effect of negative fitness is avoided by the fast rates of cultural transmission (i.e. individual shift strategy before experiencing a population decline resulting from density being over $K$). Figure XXX2 also implies that high reliance on social learning leeds to the emergence of a new basin of attraction for these unstably cycles, with small differences in initial conditions leading again to profoundly different dynamics.

%\begin{figure}
%  \centering
%      \includegraphics[width=0.5\textwidth]{./figures/figureUnstable1.pdf}
%  \caption{}
%    \label{}
%\end{figure}

%\begin{figure}
%  \centering
%      \includegraphics[width=0.5\textwidth]{./figures/figureUnstable2.pdf}
%  \caption{}
%    \label{}
%\end{figure}
